Physiological sensor based prediction of mental workload: A multimodal deep learning approach

### 2021-05-08 0.manuscript/manuscript.tex
Finishes manuscript

\documentclass[12pt]{article}

%Required packages
\usepackage[utf8]{inputenc}
\usepackage{apacite}
\usepackage{caption} 
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[textwidth=160mm, textheight=210mm, hmarginratio=1:1]{geometry}
\usepackage{geometry}
\usepackage[capposition=top]{floatrow}
\usepackage{authblk}

%Prerequisite statements
\DeclareCaptionLabelSeparator*{spaced}{\\[2ex]} %Declare the caption and title seperator spacing
\captionsetup[table]{,textfont=it,format=plain,justification=justified,
  singlelinecheck=false,labelsep=spaced,skip=2ex} %Setting up caption for table titles
 \captionsetup[figure]{,textfont=it,format=plain,justification=justified,
  singlelinecheck=false,labelsep=spaced,skip=2ex} %Setting up caption for figure titles
\captionsetup{labelfont={bf}} %Bold label font for tables
\onehalfspacing %Document line spacing
\pagestyle{myheadings} %Heading style
\graphicspath{ {./figures/} } %Defining image path

%Frontpage specifications 
\title{Predicting  Mental Workload: An Intermediate Fusion Deep Learning Approach}

\author[1]{Bart-Jan Boverhof \thanks{b.boverhof@students.uu.nl}}
\author[2]{Bernard P. Veldkamp\thanks{b.p.veldkamp@utwente.nl}}
\affil[1]{\normalsize Faculty of Behavioral, Management and Social Sciences, University of Twente}
\affil[2]{\normalsize Faculty of Social and Behavioural Sciences, Utrecht University}

\renewcommand\Authands{ and }
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
Morbi tempor congue porta. Proin semper, leo vitae faucibus dictum, metus mauris lacinia lorem, ac congue leo felis eu turpis. Sed nec nunc pellentesque, gravida eros at, porttitor ipsum. Praesent consequat urna a lacus lobortis ultrices eget ac metus. In tempus hendrerit rhoncus. Mauris dignissim turpis id sollicitudin lacinia. Praesent libero tellus, fringilla nec ullamcorper at, ultrices id nulla. Phasellus placerat a tellus a malesuada.
\end{abstract}

\hspace*{3.6mm}\textit{Keywords:} lorem, ipsum, dolor, sit amet, lectus % Keywords

\newpage
\section{Introduction}
The topic of mental workload is a widely studied phenomenon across a variety of different fields, amongst others the field of ergonomics \cite{young2015state}, human factors \cite{pretorius2007development} and cognitive neurosciences \cite{shuggi2017mental}. A commonly utilized definition of mental workload, hereafter referred to as simply "workload", is the demand placed upon one whilst one carries out a particular task. As rightfully pointed out by \citeA{de1996measurement}, the aforementioned definition is lacking, for it defines workload solely as a phenomenon external to the individual. Workload, however, requires to be recognized as a person-specific construct, for the amount of perceived workload ushered by a given task may vary substantially across individuals. \cite{de1996measurement}.

A commonly employed method for the assessment of workload is the well established NASA-Task Load Index questionnaire, hereafter referred to as "NASA-TLX". This method inquires into the amount of perceived workload,  embodying six different dimensions \cite{hart2006nasa}. Due to practical considerations such an assessment is usually conducted post-experiment, which can in certain situations be deemed suboptimal. Consider for example an experiment in which it is aimed to assess the perceived workload of a pilot in flight. An insightful  approach towards such an experiment would be to measure the degree of perceived workload during different phases of the flight, such that can be determined which specific manoeuvres tend to cause an increase in perceived workload.  However,  a questionnaire-based assessment such as the NASA-TLX can only be administrated after the flight is concluded. In such a situation, utilizing a post-experiment assessment well after the action took place is prone to generate biases. One example of such a bias is the observer-bias, dictating that participants of an experiment tend to overexaggerate the treatment effect, i.e. the amount of perceived workload in our case, when having to report it post-experiment \cite{mahtani2018catalogue}.

An alternative approach to the assessment of workload is to collect physiological bio-signals during the experiment, and utilize these inputs to predict the degree of experienced workload.  A considerable advantage of such an approach is that it renders possible to cater the method towards the individual,  hereby acknowledging the between-personal perceptual differences inherent to workload, the importance of which was discussed in the first paragraph of this paper and stressed by \citeA{de1996measurement}. But taking a physiological bio-signal-approach poses more advantages: e.g. one may utilize multiple complementary bio-signals, each stemming from a different physiological source simultaneously \cite{ramachandram2017deep}. Doing so poses the theoretical potential of yielding a rich and multifaceted assessment of a mental construct such as workload. In addition, with such an approach one may select a variety of the most suitable physiological bio-signals, hereafter referred to as "modalities",  and combine these information streams in order to yield a multifaceted prediction.

With the current research endeavor we explore the feasibility of an approach to workload assessment by means of multiple modalities simultaneously, hereafter referred to as "multimodal approach". We do so by combining data-streams stemming from three different modalities, being electroencephalography, photoplethysmography and galvanic skin response. Electroencephalography,  hereafter referred to as "EEG", is often utilized in the assessment of workload \cite{craik2019deep} \cite{berka2005evaluation} and found to be amongst the most adequately performing techniques within the field \cite{hogervorst2014combining}. Galvanic skin response, hereafter referred to as "GSR", equally so is a widely adopted approach towards workload assessment \cite{nourbakhsh2012using} \cite{zhou2015dynamic}. Lastly, heart-rate is a widely recognized indicator of workload, mostly obtained through photoplethysmography, hereafter referred to as "PPG" \cite{zhang2018evaluating} \cite{jimenez2018using}.

A deep learning approach towards modeling is adopted. A range of four different deep neural networks, hereafter referred to as "DNN's", are constructed and contrasted in performance: three of which are unimodal DNN's, i.e. networks each utilizing data from only a single modality. Thus,  one DNN solely utilizes the EEG modality,  one solely utilizes GSR and one solely utilizes PPG. The fourth network is a multimodal network combining all three modalities into one. With this research we firstly sought to explore which modalities constitute the most adequate predictors of workload in an experimental setting, and secondly whether a multimodal combination is preferable over the three simpler (but much less computationally demanding) unimodal DNN's. This assessment will be propelled by contrasting network performance, however when drawing conclusions, considerations regarding computational costs will be taken into consideration as well. All networks were constructed in Python, by utilizing the deep-learning toolbox PyTorch \cite{paszke2017automatic}.

%Method section
\newpage
\section{Data \& Methods}

\subsection{Data}
The subsequent section will provide an overview of the utilized data. Attention is placed upon the experimental setup, the description of the respondents, the utilized devices for data collection and the synchronization process.

\subsubsection{Experimental Setup}
The experimental setting in which data collection befell is the open-source spaceship video-game Empty Epsilon, in which partakers carry out tasks on a virtual spaceship bridge \cite{daid2016empty}. This experiment was instituted by the Brain Computer Interfaces testbed, hosted by the University of Twente and carried out in cooperation with Thales group Hengelo. The experiment constituted three different segments, in each of which the respondent had to carry out tasks. These tasks were designed to evoke varying degrees of workload. Each segment consisted of six small sessions, lasting roughly five to ten minutes. These sessions varied in difficulty, including two easy, two intermediate and two hard sessions per segment. Within each segment, the order in which the sessions were presented was randomized. Between every three sessions, respondents were given a short two minute break. A schematic overview of the experimental structure is depicted as Figure \ref{fig:experimental}.

After each of the 18 sessions, respondents were asked to fill out the NASA-TLX, resulting in 18 filled in questionnaires per respondent. Each questionnaire consists out of 6 questions, inquiring upon the degree to which the respondent experienced workload during the preceding session. The mean of a subset of four of these items have been utilized to label the data for training purposes.  Two items were not incorporated into the scale, for they inquired into physical demand specifically, which given the experiment took place on a computer whilst sitting behind a desk was considered to be a redundant factor. Each item, and hence the newly constructed scale, ranges from 0-20, wherein 0 reflects low workload and 20 reflects high workload.  In order to encourage numerical stability, label scores were normalized to reside in-between 0-1.

\begin{figure}
\caption{Schematic Depiction of the Experimental Setup}
\bigskip
\includegraphics[scale=0.35]{experimental_setup}
\label{fig:experimental}
\bigskip
\end{figure}

The first segment emulated a scenario in which hostile spaceships approached the respondent's spaceship. The respondent was required to quickly respond by defusing the hostiles in order to survive. The increment in difficulty caused the process of defusing hostile spaceships to become more challenging, hereby aiming to cause an increase in workload. The second segment emulated a scenario in which the respondent had to navigate their spaceship trough space, gathering as many way-points as possible. Obstacles around which the respondent had to navigate carefully, as well as hostile spaceships the respondent had to decimate, were introduced in the higher difficulty sessions. The third and final segment emulated a machine room, in which respondents had to control the power based on randomly generated requests. In the increased difficulty settings, variables that could overheat the spaceship were introduced, demanding the respondent to multi-task and aiming to increase workload as a consequence.

\subsubsection{Participants}
After having to omit 7 participants due to hardware failure, 27 respondents were included in the final analysis.  18 participants are female whereas 9 are male. The mean age and standard deviation are $\mu =26, \sigma = 10.31$ respectively. The participants are students recruited from the University of Twente, as well as several employees of Thales group Hengelo.

\subsubsection{Devices and Synchronization}
The Shimmer3 GSR+ sensor was used for both PPG and GSR measurements. This device is worn on the wrist. Signals are communicated wirelessly via Bluetooth. An ear-clip is utilized for measuring PPG. The Shimmer3 GSR+ automatically converts measured PPG signals to heart-rate. Skin conductivity, or GSR, is measured by two electrodes attached to the fingers \cite{shimmer}. EEG measurement was conducted with the Muse 2, which is a commercially offered multi-sensor headband that provides feedback on brain activity \cite{muse}. The Muse 2 headband constitutes five sensors, each monitoring a different brainwave frequency.

The Shimmer3 GSR+ measures signals on a sampling rate of 256 Hz, whereas the Muse 2 measures at a sampling rate of 220 Hz. Given that multiple devices are utilized, data streams are required to be properly synchronized. This was accomplished by means of an application called Lab-Streaming Layer, hereafter referred to as "LSL", developed by \citeA{kothe2018lab}. The three data streams stemming from the two devices were all streamed to LSL during the experiment. LSL subsequently properly synchronizes all data streams in real-time,  such that they are parallel, hence all referring to equivalent points in time.

\subsection{Related Work} 
The following section provides an overview of preceding research regarding the assessment of workload by means of physiological bio-signals, with a focus on deep learning approaches. Attention is predominantly placed upon the most feasible network architectures, in addition to a range of model optimization techniques and to the data fusion strategy for the multimodal DNN. The final part of this section will be dedicated to hyperparameter optimization.

\subsubsection{Electroencephalography (EEG)}
An overview of the complete literature on EEG applications with deep-learning was presented by \citeA{craik2019deep}, who reported a total of 16 \% of all publications to consider workload assessment specifically. The lion's share of these publications utilized either a deep belief networks or convolutional neural network, hereafter referred to as "ConvNet". One of these approaches is encouraged by the authors consequently \cite{craik2019deep}.

\citeA{tabar2016novel} proposed a hybrid of a ConvNet with a stacked auto-encoder network, hereafter referred to as "SAE network". Inputs were specified to feed into the convolutional-part with the objective of learning extracting features. The output of this part was subsequently specified to feed into the SAE part of the network,  which encompasses a stack of dense-layers, designed specifically as an input layer, six hidden layers and an output layer. A classification accuracy of 90 \% was acquired with this network \cite{tabar2016novel}.

Research by \citeA{schirrmeister2017deep} contrasted the performance of several ConvNets against the widely utilized baseline method for EEG classification, filter bank common spatial pattern, hereafter referred to as "FBCSP". A deep ConvNet, a shallow ConvNet, a deep-shallow hybrid ConvNet and a residual ConvNet were contrasted with an FBCSP. Both the deep and shallow ConvNets were found to reach at least similar, and in some regards better classification results as compared with the FBCSP baseline approach. Altogether, a deep ConvNet with four convolutional-max-pooling blocks was found to perform best, exhibiting an accuracy of 92.4\% \cite{schirrmeister2017deep}.

\subsubsection{Galvanic Skin Response (GSR)}
\citeA{sun2019hybrid} explored the most suitable approach to the classification of several emotional states with GSR. Various models were explored, amongst others a support vector machine, a ConvNet and a long-short-term-memory, hereafter referred to as "LSTM", network. Additionally, the feasibility of a hybrid DNN, combining both the ConvNet and LSTM approaches, was explored. This aforementioned hybrid model was found to perform best, exhibiting an accuracy of 74\% \cite{sun2019hybrid}.

\citeA{dolmans2020perceived} took, amongst other approaches to workload prediction,  a GSR approach, and designed a variant based on the previously delineated CovNet-LSTM structure. The performance of this model was contrasted with a network consisting solely of fully connected dense layers. Conform with findings by  \citeA{sun2019hybrid}, the hybrid model was found to perform best, displaying an absolute difference between predicted and true label of 0.197 (scaled on 0-1). The model architecture as utilized deployed two convolutional max-pooling blocks and two LSTM layers \cite{dolmans2020perceived}.

\subsubsection{Photoplethysmography (PPG)}
Research by \citeA{biswas2019cornet} explored a deep learning approach towards PPG classification, with the objective to perform both bio-metric identification and obtain heart rate information. Exceptional results were realized with a DNN, attaining an average accuracy of 96\% \cite{biswas2019cornet}. This performance was realized with a ConvNet-LSTM hybrid, incorporating two convolutional max-pooling blocks followed by two LSTM layers.

\subsubsection{Multimodal Fusion Strategy}  
When conducting a multimodal deep learning approach,  information streams stemming from the different modalities are required to be combined, i.e. "fused", at a certain point in the network in order to ultimately result in a single prediction of workload. Fusion can be done conforming several strategies. Three strategies as proposed by \citeA{ramachandram2017deep} have been considered.

Early, or data-level, fusion constitutes an approach that fuses data sources before being fed into the network. Early fusing is usually proven to be quite challenging, residing in the fact that data streams stemming from different modalities often differ in their dimensionality and sampling rate. In addition, when taking an early fusion approach, the oversimplified assumption of conditional independence is made implicitly. This assumption is unrealistic in nature, for data stemming from different modalities are expected to be correlated in practice \cite{ramachandram2017deep}.

Late, or decision-level, fusion refers to the process of aggregating the decisions of multiple separate networks, each applied towards every modality separately. In case the data sources stemming from the various modalities are either correlated or ultimately differ in their dimensionality, late fusion is often a more feasible approach as opposed to early fusion \cite{ramachandram2017deep}.

Lastly, intermediate fusion is the most widely employed fusion strategy for multimodal deep-learning problems. Data streams are usually fused by a concatenation layer, joining the outputs of the separately defined network parts of each modality. This results in a single joint deep neural network. Several "higher-order layers" are usually defined in between the concatenation layer and the ultimate classification. The depth of the fusion, i.e. the specified number of higher-order layers, can be chosen conform to the specific circumstances, posing intermediate fusion to be the most flexible and therefore the most widely adopted fusion strategy \cite{ramachandram2017deep}.

Indeed, when consulting the literature, one is prone to conclude that intermediate fusion strategies are the most prevailing for multimodal deep learning approaches. When taking an intermediate fusion approach, the higher-order part of the network needs to be designed and established, for which several previous research endeavors are considered.  \citeA{rastgoo2019automatic} utilized a multimodal ConvNet approach, and fused the modalities by concatenation, followed with two LTSM layers and two dense layers. A simpler approach is adopted by \citeA{han2020classification}, who utilized an intermediate fusion approach solely consisting of several fully connected dense layers. Lastly, \citeA{dolmans2020perceived} took a relatively deep intermediate fusion approach, consisting of two dense layers, two convolutional layers followed by another two dense layers.

\subsubsection{Model Optimization Strategies}
The technique of batch normalization was initialized by \citeA{ioffe2015batch}, and is widely utilized in DNN's with the objective of stabilization. Especially ConvNets benefit from this technique. It is beneficial to incorporate a batch normalization layer subsequent to a convolutional layer, but before feeding into the activation function \cite{ioffe2015batch}.  All previously delineated networks architectures utilize batch normalization layers, in most cases specified subsequent to each convolutional layer \cite{dolmans2020perceived} \cite{tabar2016novel} \cite{schirrmeister2017deep} \cite{biswas2019cornet} \cite{sun2019hybrid}.

Pooling layers are commonly employed in ConvNets, usually succeeding a convolutional layer with the purpose of reducing dimensionality. The objective of such layers are to down-sample features into a more compact space, hereby only retaining indispensable information. For a more extensive elaboration on pooling, please consult \citeA{lecun2015deep}. All previously delineated networks architectures utilize max-pooling layers, usually specified subsequent to the activation layer \cite{dolmans2020perceived} \cite{tabar2016novel} \cite{schirrmeister2017deep} \cite{biswas2019cornet} \cite{sun2019hybrid}.

\subsubsection{Hyper Parameter Optimization}
Hyper parameter optimization, hereafter referred to as "HPO", is a technique that can be utilized to optimize network hyper-parameters, such as learning rate and dropout rate.  In theory, one could optimize the entire DNN architecture, including the number of neurons/filters in (convolutional) layers, the number of layers in general, whether to use certain layers etc. There is, however, a strong relationship between the amount of parameters to be optimized and the computational resources required to do so, in where many parameters to optimize tends to inflate computational costs by a substantial amount. Substantial advancements in DNN performance have been attained by utilizing HPO, especially for ConvNets \cite{bergstra2012random}. The Optuna toolbox provides a method for creating a parameter search space, from which values for the hyper-parameters can be sampled and optimization can be performed \cite{akiba2019optuna}. 
\bigskip

>>>>

\documentclass[fleqn,11pt]{paper}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{apacite}
\renewcommand{\doiprefix}{}
\newcommand{\doi}[1]{https://doi.org/#1}

\title{Physiological sensor based prediction of mental workload: A multimodal deep learning approach}

\author[1]{Bart-Jan Boverhof \thanks{b.boverhof@students.uu.nl}}
\affil[1]{\normalsize Faculty of Social and Behavioural Sciences, Utrecht University}

\begin{abstract}
The assessment of mental workload by means of physiological sensors has gained in popularity in recent years. We utilized three physiological sensors from which mental workload was predicted, being electroencephalography, photoplethysmography and galvanic skin response. Deep learning poses a promising, yet novel, approach towards the modeling of such sensory data. Four deep neural network architectures were created and contrasted in their performance. Three out of four were specified to predict workload based on on a single of the aforementioned physiological sensors, whilst the fourth architecture was a joint neural network combining all physiological signals into a single model. In this study we aimed to identify 1) which physiological sensor constitutes the most adequate predictor of mental workload, and 2) whether a joint approach utilizing multiple sensors is preferable over the approaches utilizing just one. Altogether the electroencephalography-only architecture was found to perform best, closely followed by the galvanic skin response-only architecture,  both displaying a mean absolute error of 0.110 and 0.119 (scaled on 0-1) respectively. Against expectations, the joint approach was outperformed by both aforementioned architectures, displaying a mean absolute error of 0.128 (equally so scaled on 0-1). The photoplethysmography-only architecture was found to perform worst, displaying a mean absolute error of 0.136. In the discussion section we presented several explanations that help rationalize this anomaly, one being due to an imperfect model architecture for the joint neural network, and another being due to overfitting resulting from a limited amount of data. Regardless of the unexpected character of this result, we still recognize that differences in performance were of rather modest size: all four neural network architectures performed fairly adequately. 
\end{abstract}

\begin{document}
\begin{titlepage}
\vspace{10mm}

\begin{large}
{\setlength{\parindent}{0cm}
Methodology and Statistics for the Behavioural, Biomedical and Social Sciences
}
\newline Utrecht University, the Netherlands
\vspace{18mm}
\newline MSc Thesis Bart-Jan Boverhof (6000142)
\newline Physiological sensor based prediction of mental workload: A multimodal deep learning approach
\newline May 2021
\vspace{18mm}
\newline Supervisor:
\newline Prof. Dr. Ir. Bernard Veldkamp
\vspace{10mm}
\newline Second grader:
\newline Prof. Dr. Ir René Eijkemans
\vspace{10mm}
\newline Preferred journal of publication: Frontiers in Neuroscience
\end{large}
\end{titlepage}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

\section{Introduction}
The topic of mental workload is a widely studied phenomenon across a variety of different fields, amongst others the field of ergonomics \cite{young2015state}, human factors \cite{pretorius2007development} and cognitive neurosciences \cite{shuggi2017mental}. A commonly utilized definition of mental workload, hereafter referred to as simply "workload", is the demand placed upon one whilst carrying out a particular task. As rightfully pointed out by \citeA{de1996measurement}, the aforementioned definition is incomplete, for it defines workload solely as a phenomenon external to the individual. Workload, however, requires to be recognized as a person-specific construct, for the amount of perceived workload ushered by a given task may vary substantially between individuals \cite{de1996measurement}.

A commonly employed method for the assessment of workload in an experimental setting is the well established NASA-Task Load Index questionnaire, hereafter referred to as \enquote{NASA-TLX}. This method embodies several dimensions, jointly inquiring into the amount of perceived workload \cite{hart2006nasa}. However, due to the ever-increasing availability in computing power characterizing recent times, a novel approach to workload assessment has gained in popularity. This promising approach constitutes the use of physiological sensors, worn throughout the entire duration of the experiment. An approach like this poses the advantage that the development of perceived workload may be monitored throughout the entire experiment, enabling the possibility to precisely monitor the impact of a specific task on a participant's level of workload. Another advantage is that the method may be personalized, which could simply be done by creating and training an individual model for each participant. By doing this, the person-specific character that is inherent to workload is acknowledged, the importance of which was stressed in the opening paragraph of this paper.

With the current research we explore the feasibility of an approach towards workload assessment by means of various physiological sensors, hereafter referred to as \enquote{modalities}. The three utilized modalities are electroencephalography (EEG), photoplethysmography (PPG) and galvanic skin response (GSR). EEG is a technique that monitors electrical activity in the brain, and is often utilized in the assessment of workload \cite{berka2005evaluation, craik2019deep}. EEG is found to be amongst the most adequately performing techniques for this task \cite{hogervorst2014combining}. GSR is a technique that measures electrical conductance of the skin, and equally so is a widely adopted method for workload assessment \cite{nourbakhsh2012using, zhou2015dynamic}. Lastly, also heart-rate is a widely utilized indicator of workload, mostly obtained through PPG \cite{zhang2018evaluating,jimenez2018using}.

In this project we opted for a deep learning approach towards workload prediction. A range of four different deep neural network architectures, hereafter referred to as \enquote{DNN's} or simply \enquote{networks},  are constructed and contrasted in their performance. Among these are three unimodal DNN's, i.e. networks each utilizing only a single modality. Thus,  one network solely utilizes the EEG modality,  one solely utilizes the GSR modality and one solely utilizes the PPG modality. The fourth is a multimodal network, i.e. combining all three aforementioned modalities into a single architecture. The main advantage of taking a deep learning approach specifically resides in its flexibility. Fitting a model that combines multiple modalities can be done in different, relatively straightforward, ways when taking a deep learning approach \cite{ramachandram2017deep}. We expect the multimodal approach to results in better performance as compared with the unimodal approaches, for predictions are based on richer composite of information. This expectation is indeed in line with previous research, which reported a multimodal approach to perform superiorly \cite{dolmans2020perceived, han2020classification,rastgoo2019automatic, yin2017recognition}.

With this research we 1) seek to explore which individual modality constitutes the most adequate predictor of workload when taking a deep learning approach, and 2) whether a multimodal network architecture is preferable over the three simpler (but much less computationally demanding) unimodal architectures. This comparison will mainly be propelled by contrasting network performance, however when drawing conclusions, considerations regarding computational costs will also be taken into consideration. With this research we aim to deepen the knowledge of taking a deep learning approach towards workload prediction with physiological sensors in an experimental setting, which as of yet, is still a novel and fairly understudied topic.

%Method section
\vspace{10mm}
\section{Data \& Methods}
\subsection{Related Work} \label{sec:related}
The following section provides an overview of preceding research on the assessment of workload by means of various physiological sensors, with a focus on deep learning approaches. Attention is predominantly placed upon network architectures and hyperparameter optimization. In this section we will solely describe previous work: an in depth description of our adopted modeling approach follows in Section \ref{section:DNN}.

\subsubsection*{EEG Network Architecture}
An overview of the complete literature EEG applications with deep learning was presented by \citeA{craik2019deep}, who reported 16\% of all publications to constitute workload assessment. The lion's share of these publications utilized either a deep belief network or convolutional neural network (ConvNet). \citeA{tabar2016novel} proposed a ConvNet and stacked auto-encoder (SAE) hybrid approach towards workload classification with EEG. Their proposed network constituted two different parts. Firstly, raw EEG data was fed into a single convolutional layer, with the objective of extracting features. Secondly, the the extracted features fed into the SAE part of the network,  which encompassed a stack of dense-layers designed to ultimately result in a classification of workload.  An accuracy of 90\% was acquired with this architecture \cite{tabar2016novel}. A similar approach was proposed by \citeA{schirrmeister2017deep}, who contrasted different DNN architectures in their ability to adequately decode EEG signals. Several ConvNet architectures were contrasted against the baseline method for EEG decoding, filter bank common spatial pattern (FBCSP). A deep ConvNet, a shallow ConvNet, a deep-shallow hybrid ConvNet and a residual ConvNet were contrasted with the FBCSP baseline method. Both the deep and shallow ConvNets were found to reach at least similar, and in some regards better classification results as compared with the FBCSP baseline approach. Altogether, a deep ConvNet with four convolutional-max-pooling blocks was found to perform best, exhibiting a classification accuracy of 92.4\% \cite{schirrmeister2017deep}.

\subsubsection*{GSR Network Architecture}
\citeA{sun2019hybrid} explored various deep-learning approaches towards the classification of emotional states with GSR. Despite that this application differs from workload prediction, inspiration may still be drawn from their utilized architecture. Various models were explored, amongst others a support vector machine, a ConvNet and a long-short-term-memory (LSTM) network. In addition to these, a hybrid ConvNet-LSTM approach was explored as well. This hybrid approach constituted four convolutional blocks followed by a single LSTM layer, and was found to perform best with an accuracy of 74\% \cite{sun2019hybrid}. \citeA{dolmans2020perceived} contrasted several physiological signals in their ability to predict workload, each of which by taking a deep learning approach. One of these modalities constituted GSR, for which a ConvNet-LSTM hybrid network architecture was utilized to predict workload. The performance of this DNN was contrasted with a network consisting solely of fully connected dense layers. Conform with the previously mentioned findings by  \citeA{sun2019hybrid}, the ConvNet-LSTM hybrid architecture was found to perform best, displaying an absolute difference between predicted and true label of 0.180 (scaled on 0-1). This network architecture incorporated two convolutional blocks, followed by two LSTM layers \cite{dolmans2020perceived}.

\subsubsection*{PPG Network Architecture}
Research by \citeA{biswas2019cornet} explored a PPG deep learning approach, in which their objective was to perform both bio-metric identification and obtain heart rate information. Again, despite that this research doesn't constitute workload prediction or classification, inspiration may still be drawn from the employed network architecture. An exceptional performance of 96\% accuracy was realized with a ConvNet-LSTM hybrid, incorporating two convolutional max-pooling blocks followed by two LSTM layers \cite{biswas2019cornet}. \citeA{dolmans2020perceived} also explored a deep-learning PPG approach to workload prediction.  A ConvNet constituting two convolutional blocks was found to perform best out of all contrasted PPG architectures, displaying an absolute difference between predicted and true label of 0.1969 (scaled on 0-1).

\subsubsection*{Multimodal Fusion Strategy}  
When creating a DNN that combines multiple modalities into a single network,  the information streams that stem from the different subparts of the network (each corresponding to one of the modalities) are required to be combined, i.e. \enquote{fused}, at a given point in the network. This is necessary in order to ultimately end up with a single prediction of workload. Fusion can be done conforming several strategies, out of which three strategies as proposed by \citeA{ramachandram2017deep} are considered.

Firstly, early fusion constitutes an approach that fuses data sources before being fed into the network. Early fusing usually proves to be quite challenging, residing in the fact that data streams stemming from different modalities often differ in their dimensionality and sampling rate. In addition, when taking an early fusion approach, the assumption of conditional independence is made implicitly. This assumption is oftentimes unrealistic in practice, for data stemming from different modalities are expected to be correlated \cite{ramachandram2017deep}. Secondly, late fusion refers to the process of aggregating the decisions of multiple separate networks, each constructed for every modality separately. When the data sources stemming from the various modalities happen to be either correlated or different in their dimensionality, late fusion is oftentimes a more feasible approach as opposed to early fusion \cite{ramachandram2017deep}. Lastly, intermediate fusion is the most widely employed fusion strategy for multimodal networks. Data streams are usually fused by a concatenation layer, joining the outputs of the separately defined network parts of each modality. This results in a single joint deep neural network. Several higher-order layers are usually defined in between the concatenation layer and the ultimate prediction layer. The depth of the fusion, i.e. the specified number of higher-order layers, may be chosen by the researcher, posing intermediate fusion to be the most flexible and therefore the most widely adopted fusion strategy \cite{ramachandram2017deep}.

Indeed when consulting the literature, it becomes apparent that intermediate fusion is the most prevailing strategy for multimodal DNN architectures. As mentioned before, when taking an intermediate fusion approach, the higher-order part of the network is required to be designed by the researcher. \citeA{rastgoo2019automatic} utilized a multimodal ConvNet approach, and fused the modalities by concatenation, followed with two LTSM layers and two dense layers. A simpler approach is adopted by \citeA{han2020classification}, who utilized an intermediate fusion approach solely consisting of several fully connected dense layers. Lastly, \citeA{dolmans2020perceived} took a relatively deep intermediate fusion approach, consisting of two dense layers, two convolutional layers followed by another two dense layers.

\subsubsection*{Auxiliary Layers in a DNN Architecture}
At the core of a DNN are the aforementioned types of layers, such as convolutional, LSTM and dense layers. Such layers are often accompanied with auxiliary layers in order to ensure network stability. The first type of layers often used in network architectures are batch normalization layers, which have the objective improving network stability. It is beneficial to incorporate a batch normalization layer subsequent to a convolutional layer, but before feeding into the activation layer \cite{ioffe2015batch}.  For a deep dive into batch normalization, we refer the reader to \citeA{ioffe2015batch}. All previously considered network architectures made use of batch normalization layers \cite{biswas2019cornet, dolmans2020perceived, schirrmeister2017deep, sun2019hybrid, tabar2016novel}.

Pooling layers are commonly employed in ConvNets, usually succeeding a convolutional layer with the purpose of reducing dimensionality. The objective of such layers are to down-sample features into a more compact space, hereby only retaining essential information, and thus omitting redundant information. For a more extensive elaboration on pooling we refer the reader to \citeA{lecun2015deep}. All previously considered network architectures utilize max-pooling layers, usually specified subsequent to the activation layer \cite{biswas2019cornet, dolmans2020perceived, schirrmeister2017deep, sun2019hybrid, tabar2016novel}.

The incorporation of dropout layers was initially proposed by \citeA{srivastava2014dropout}. Dropout layers drop a randomly-selected number of neurons and their connections throughout the training process. More specifically, within each full training run through the data, i.e. each \enquote{epoch}, a different randomly selected set of neurons are fixed to be inactive. Doing so helps to reduce overfitting, particularly for complex network architectures \cite{srivastava2014dropout}.  Most of the previously considered network architectures make use of dropout layers \cite{biswas2019cornet, dolmans2020perceived, schirrmeister2017deep, sun2019hybrid}.

\subsubsection*{Hyperparameter Optimization}
Hyperparameter optimization (HPO) is a technique that optimizes network hyperparameters, such as for example learning rate and dropout rate. Substantial advancements in DNN performance have been attained by utilizing HPO, especially in the case of ConvNets \cite{bergstra2012random}. In theory, one could optimize the entire DNN architecture, including the number of neurons/filters in dense/convolutional layers, the number of layers in general, whether to use certain layers, etc. There is, however, a strong relationship between the amount of parameters to be optimized and the computational resources required to do so: optimizing many hyperparameters will inflate computational costs substantially. The Optuna toolbox provides a method for creating a parameter search space, from which values for the hyperparameters may be sampled, and optimization can be performed \cite{akiba2019optuna}.

\subsection{Data} \label{sec:data}
The subsequent section provides a description of the utilized data and its collection process. Attention is placed upon the experimental setup, the description of the respondents, the utilized devices for data collection, the synchronization process and the data preprocessing approach.

\subsubsection*{Experimental Setup \& Data Labeling Strategy}
The experimental setting in which our data was collected is the open-source spaceship video-game Empty Epsilon, in which partakers carried out tasks on a virtual spaceship bridge \cite{daid2016empty}. This experiment was instituted by the Brain Computer Interfaces Testbed \footnote{https://bmslab.utwente.nl/} hosted by the University of Twente and carried out in cooperation with Thales Group\footnote{https://www.thalesgroup.com/en/countries/europe/netherlands} Hengelo, the Netherlands. The experiment was divided into a total of 15 tasks, differing in difficulty and hence designed to evoke varying degrees of workload. The total duration of the experiment was roughly three hours.

Respondents were asked to fill out the NASA-TLX questionnaire after each of the fifteen task, inquiring into the amount workload perceived during the respective task. Each questionnaire consisted out of six items. The mean of a subset of four of these items were utilized for labeling of the data. Two items were not incorporated into the scale, for they inquired into the perceived physical demands. Since the experiment took place on a computer whilst participants sat behind a desk, we chose to not incorporate such items. Each item, and hence the newly constructed scale, ranges from 0-20, wherein 0 reflects the lowest possible level workload and 20 reflects the highest possible level workload.  In order to encourage numerical stability, label scores were normalized to reside in-between 0-1.

\subsubsection*{Participants}
After having to omit seven participants due to hardware failure, 27 respondents remained in the final analysis.  18 participants were female whereas nine were male. The average age and its standard deviation were $\mu =26$ and $\sigma = 10.31$ respectively. The participants were students recruited from the University of Twente situated in the Netherlands, as well as several employees of Thales Group Hengelo, equally so situated in the Netherlands.

\subsubsection*{Devices and Synchronization}
The Shimmer3 GSR+ sensor was used for both PPG and GSR measurements. This device was worn on the wrist, whilst signals were communicated wirelessly via Bluetooth. An ear-clip was utilized for measuring PPG, which was automatically converted to heart-rate. Skin conductivity, or GSR, was measured by two electrodes attached to the fingers. EEG measurement was conducted with the Muse 2 multi-sensor headband, equally so whilst communicating signals wirelessly via Bluetooth. The Shimmer3 GSR+ measured on a sampling rate of 256 Hz, whereas the Muse 2 measured at a sampling rate of 220 Hz.

Given that multiple devices were utilized, data streams were required to be properly synchronized. This was accomplished by means of an application called Lab-Streaming Layer, hereafter referred to as "LSL", developed by \citeA{kothe2018lab}. The three data streams stemming from the two devices were all streamed to LSL during the experiment. LSL subsequently properly synchronizes all data streams in real-time,  such that they are parallel, hence all referring to equivalent points in time.

\subsubsection*{Data Preprocessing \& Partitioning}
For each participant, the EEG, PPG and GSR data streams were cut into windows of eight seconds, resulting in an average of roughly 351 windows per respondent. 80\% of all windows have been used for training, 10\% for validation within the training process and 10\% for the testing of network performance. The partitioning of the windows into training, validation and testing sets was done systematically by selecting every \textit{n-th} window, such that windows over the entire duration of the experiment were represented equally in each partition.  Exactly the same partitions were utilized in the training, validation and testing procedure across the four architectures to ensure a fair comparison.
